---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
```

# Load the data

```{r}
# Load the data
data_A <- read_table("Data_Task/Data/sumstats_trait_A.txt")
data_A |>
    filter(SNP == "rs1110052")
```

# check the data and clean or fix 

## get a full data set

### missing value in position

```{r}
rows_to_change <- which(data_A$BPOS %in% c("T", "C", "G", "A"))
length(rows_to_change)

data_A_new <-
    data_A[rows_to_change, ] |>
    mutate(
        info = NCHROBS,
        NCHROBS = z,
        z = beta_hat,
        beta_hat = N,
        N = MAF,
        MAF = A2,
        A2 = A1,
        A1 = BPOS,
        BPOS = NA
    )

data_A <- rbind(data_A[-rows_to_change, ], data_A_new)
```

### missing value in NCHROBS


We find that the individuals who lose its  NCHROBS also lose its z and info, so we find that we should replace NCHROBS with beta_hat.

```{r}
rows_to_change <- which(is.na(data_A$NCHROBS))
length(rows_to_change)

data_A_new <-
    data_A[rows_to_change, ] |>
    mutate(
        NCHROBS = beta_hat,
        beta_hat = NA
    )

data_A <- rbind(data_A[-rows_to_change, ], data_A_new) |>
    mutate(MAF = as.numeric(MAF))
```

Now we have the full data set and can have a further check on each variable.

## check each variable

### check A1, A2 and MAF

```{r}
data_filtered <- data_A |>
    filter(
        A1 %in% c("A", "C", "G", "T") &
            A2 %in% c("A", "C", "G", "T") &
            MAF > 0 & MAF < 0.5 &
            A1 != A2
    )
```

### check the accuracy and significance of z

Cause we use a two-sided test, so we assume a 95% significance level
and the z value should be out of -1.96 and 1.96.

```{r}
data_filtered <- data_filtered |>
    filter((z < -1.96 | z > 1.96)     
           & info > 0.90
           & NCHROBS > 800000
           & N > 70000)
```

Since GWAS is essentially similar to simple linear regression (SLR), the beta and z values should have the same sign.

```{r}
data_filtered <- data_filtered |>
    filter((beta_hat > 0 & z > 0) | (beta_hat < 0 & z < 0))
```

### position information: SNP, CHR, BPOS

```{r}
# snp duplicate？

any(duplicated(data_filtered$SNP))
duplicated_values <- data_filtered $SNP[duplicated(data_filtered $SNP)]
data_filtered  |>
    filter(SNP == duplicated_values[1])


any(!(data_A$CHR == 22))
```